---
title: "Final Assignment"
author: "Valeria Contreras"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: false
    df_print: paged
---

```{=html}
<style>
body {
text-align: justify}
</style>
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, message=FALSE, warning=FALSE, knitr.purl.inline = TRUE )
```

# Final Assignment of Survey Research II

### Libraries

```{r}
# LOADING REQUIRED LIBRARIES

library(tidyverse)
library(dplyr)
library(forcats)
library(gmodels)
library(haven)
```

## DATA COLLECTION AND WRANGLING

We load the original given dataset from the Special Eurobarometer 493: Discrimination
in the EU. This survey includes responses regarding societal attitudes
and legal challenges among 27438 respondents across 28 countries between 09-
21/05/2019.

Based on previous literature on this topic (check the narrative doc for more information),
we select and initial set of independent variables that, together with the dependent 
variable q19 ("Do you think that transgender persons should be able to change their 
civil documents to match their inner gender identity?"), will form the individual level
factors. These independent variables are:

  - Generic individual aspects:
    - ORIGIN: country and isocntry
    - GENDER: d10 (then: male)
    - AGE: d11 (then: age_cat)
    - **EDUCATION: (d8r1) ??????????????????????????**
    - RELIGION: sd3 (then: religion and religion_group)
    - GENERAL POLITICAL ATTITUDES:
    - POLITICAL STANCE: d1r1 (then: left_right)
    
  - LGTBI related aspects at individual level:
    - CONTACT: 
      - sd1_4 (gay lesbian or bisexual)
      - sd1_7 (transgender/transexual)
      - sd1_8 (intersext)
    - DISCRIMINATION PERCEIVED (LGTBI): qc1 (then: discrimination)
    - ATTITUDES TOWARDS THE LGTBI COMMUNITY:
    - EXPERIENCES OF DISCRIMINATION (LGTBI):
    - LAWS & POLICIES:
    
  - Variables at country level:
    - Unemployment rate
    - Religion
    - Rainbow (LGBTI+)
    - GDP

# Original Data

```{r}
Disc <- read_dta("ZA7575.dta")

Disc_clean <- Disc %>% 
  select(isocntry, country, d10, d11, d8r1, d1r1, sd3, sd1_4, sd1_7, sd1_8, qc19, qc20, d70, d15a_r1)
```

## Encoding: the individual perspective

Now, we need to factorize the values of the independent variables and recode the "refusal" and "DK" answers as NAs.

### Country

Encoding the country name:

```{r}
#Put the complete name of the countries. 
Disc_clean <- Disc_clean |> mutate(country = case_when(
  isocntry== "BE" ~ "Belgium",
  isocntry== "DK" ~ "Denmark",
  isocntry== "GR" ~ "Greece",
  isocntry== "ES" ~ "Spain",
  isocntry== "FI" ~"Finland",
  isocntry== "FR" ~ "France",
  isocntry== "IE" ~ "Ireland",
  isocntry== "IT" ~ "Italy",
  isocntry== "LU" ~ "Luxembourg",
  isocntry== "NL" ~ "Netherlands",
  isocntry== "AT" ~ "Austria",
  isocntry== "PT" ~ "Portugal",
  isocntry== "SE" ~ "Sweden",
  isocntry %in% c("DE-W", "DE-E")  ~ "Germany",
  isocntry== "GB" ~ "United Kingdom",
  isocntry== "BG" ~ "Bulgaria",
  isocntry== "CY" ~ "Cyprus",
  isocntry== "CZ" ~ "Czech Republic",
  isocntry== "EE" ~ "Estonia",
  isocntry== "HU" ~ "Hungary",
  isocntry== "LV" ~ "Latvia",
  isocntry== "LT" ~ "Lithuania",
  isocntry== "MT" ~ "Malta",
  isocntry== "PL" ~ "Poland",
  isocntry== "RO" ~ "Romania",
  isocntry== "SK" ~ "Slovakia",
  isocntry== "SI" ~ "Slovenia",
  isocntry== "HR" ~ "Croatia",
  TRUE ~ NA_character_))

print(Disc_clean)
```

### Encoding **dependent** variable
### qc19

Convert numeric answers from qc19 to character answers

```{r}
Disc_clean <- Disc_clean |> mutate(change_docs = case_when(
  qc19 == 1 ~ "Yes",
  qc19 == 2 ~ "No", 
  qc19 == 3 ~ "DK", 
  TRUE ~ NA_character_))

Disc_clean$change_docs <- factor(Disc_clean$change_docs, levels=c("Yes","No","DK"))

with(Disc_clean, table(country, change_docs))
```

### Encoding **independent variables**
### 1. Gender (Indepdent Variable)

Encoding gender:

```{r}
# Encoding gender (Original-> 1: Man 2: Woman)
Disc_clean <- Disc_clean %>%
  mutate(male = car::recode(d10, "2=0"),
    male = factor(male, levels = c(0, 1), labels = c("female", "male")))

Disc_clean$male <- ifelse(Disc_clean$male == "female", 0, 
                                 ifelse(Disc_clean$male== "male", 1, NA))

# Distribution of gender surveyed per country
table(Disc_clean$country, Disc_clean$male)
```

### 2. Age (Indepdent Variable)

Encoding age:

```{r}
# Encoding age (Original-> number, 99: refusal)
Disc_clean <- Disc_clean |>
  mutate(
    d11 = replace(d11, d11 > 98, NA) # 999 and other values above 
  )

Disc_clean <- Disc_clean |>
  mutate(
    age_cat = cut(d11, breaks = c(0, 19, 29, 44, 64, 98),
                  labels = c("<20", "20-29", "30-44", "45-64", ">=65"))
  )

Disc_clean$age_cat <- factor(Disc_clean$age_cat,
                                     levels = c("<20", "20-29", "30-44", "45-64", ">=65"),
                                     labels = c(0, 1, 2, 3, 4))


# Distribution of age categories by country
with(Disc_clean, table(country, age_cat))
```


### 3. Political stance: left-center-right (Independent Variable)

Encoding political stance (left-right):

1	(1 - 4) Left			
2	(5 - 6) Centre			
3	(7 -10) Right			
9	DK/Refusal

```{r}
Disc_clean <- Disc_clean |>
  mutate(
    left_right = car::recode(d1r1, "9=NA"),
    left_right = factor(left_right, labels = c("Left", "Center", "Right"))
  )

Disc_clean$left_right <- factor(Disc_clean$left_right,
                                     levels = c("Left", "Center", "Right"),
                                     labels = c(0, 1, 2))


# Distribution of political stance categories by country
with(Disc_clean, table(country, left_right))
```

### 4-1. Religion (Independent Variable) - Religion orientation

Encoding religion variable. Original:
1	Catholic			
2	Orthodox Christian			
3	Protestant			
4	Other Christian			
5	Jewish			
6	Muslim - Shia			
7	Muslim - Sunni			
8	Other Muslim			
9	Sikh			
10	Buddhist
11	Hindu			
12	Atheist			
13	Non believer or agnostic			
14	Other			
15	Refusal (SPONTANEOUS)			
16	DK

```{r}
# Encoding religions (Original-> 1-14: Religions; 15: Refusal; 16: )
Disc_clean <- Disc_clean %>% 
  mutate(
    religion = car::recode(sd3, "15=NA; 16=NA"),
    religion = factor(religion, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14),
                      labels= c("Catholic", "Orthodox Christian", "Protestant",
                                "Other Christian", "Jewish","Muslim - Shia",
                                "Muslim - Sunni","Other Muslim", "Sikh", "Buddhist",
                                "Hindu", "Atheist","Agnostic","Other")))


Disc_clean$religion <- factor(Disc_clean$religion,
                                     levels = c("Catholic", "Orthodox Christian", "Protestant",
                                "Other Christian", "Jewish","Muslim - Shia",
                                "Muslim - Sunni","Other Muslim", "Sikh", "Buddhist",
                                "Hindu", "Atheist","Agnostic","Other"),
                                     labels = c(0,1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))

with(Disc_clean, table(country, religion))
```

### 4-2. Religion (Independent Variable) - Religious or not

We also create another factor variable to capture in general terms if the person is religious or not:

```{r}
# Recoding religion valiable in the following groups: Religious: 1, Atheist: 2, Agnostic: 3, Other: 4)
Disc_clean <- Disc_clean |>
  mutate(
    religion_group = case_when(
      sd3 %in% c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11) ~ "1",
      sd3 %in% c(12) ~ "2",
      sd3 %in% c(13) ~ "3",
      sd3 %in% c(14) ~ "4",
      TRUE ~ NA_character_
    )
  ) |>
  mutate(
    religion_group = fct_recode(religion_group,
                            "Religious" = "1",
                            "Atheist" = "2",
                            "Agnostic" = "3",
                            "Other" = "4")
  )

Disc_clean$religion_group <- factor(Disc_clean$religion_group,
                                     levels = c("Religious", "Atheist", "Agnostic","Other"),
                                     labels = c(0,1, 2, 3))

with(Disc_clean, table(country, religion_group))
```

### 4-3. Religion (Independent Variable) - Contact with LGBT community

Original question: "Do you have friends or acquaintances who are…?" Variables:
sd1_4 CONTACT: gay lesbian or bisexual
sd1_7 CONTACT: transgender/transexual
sd1_8 CONTACT: intersext

```{r}
# Contact: gay, lesbian, bisexual
Disc_clean <- Disc_clean %>% 
  mutate(contact_gay = car::recode(sd1_4, "2=0; 3=NA; 4=NA"),
         contact_gay = factor(contact_gay, levels = c(0, 1), labels= c("No", "Yes"))
  )

Disc_clean$contact_gay <- factor(Disc_clean$contact_gay,
                                     levels = c("No", "Yes"),
                                     labels = c(0,1))

# Contact: transgender, transexual
Disc_clean <- Disc_clean %>% 
  mutate(contact_trans = car::recode(sd1_7, "2=0; 3=NA; 4=NA"),
         contact_trans = factor(contact_trans, levels = c(0, 1), labels= c("No", "Yes"))
  )

Disc_clean$contact_trans <- factor(Disc_clean$contact_trans,
                                     levels = c("No", "Yes"),
                                     labels = c(0,1))
 
Disc_clean <- Disc_clean %>% 
  mutate(contact_inter = car::recode(sd1_8, "2=0; 3=NA; 4=NA"),
         contact_inter = factor(contact_inter, levels = c(0, 1), labels= c("No", "Yes"))
  )

Disc_clean$contact_inter <- factor(Disc_clean$contact_inter,
                                     levels = c("No", "Yes"),
                                     labels = c(0,1))

# sd1_4
with(Disc_clean, table(country, contact_gay))

# sd1_7
with(Disc_clean, table(country, contact_trans))

# sd1_8
with(Disc_clean, table(country, contact_inter))
```

### 5. Life satisfaction (Independent variable) 

```{r}
Disc_clean <- Disc_clean %>% mutate(Life_satisf = case_when(
  d70 == 1 ~ "Very_satisfied",
  d70 == 2 ~ "Fairly_satisfied", 
  d70 == 3 ~ "Not_very_satisfied",
  d70== 4 ~ "Not_at_all_satisfied",
  d70== 5 ~ "DK",
  TRUE ~ NA_character_))  

Disc_clean$Life_satisf <- factor(Disc_clean$Life_satisf,
                                     levels = c("Very_satisfied", "Fairly_satisfied", "Not_very_satisfied",
                                                "Not_at_all_satisfied", "DK"),
                                     labels = c(0,1, 2, 3, NA))
```

### 6. Unemployement (Independent variable)

```{r}
Disc_clean <- Disc_clean %>% 
  mutate(unemployed = case_when(
    d15a_r1 == 1 ~ "No",
    d15a_r1 == 2 ~ "No", 
    d15a_r1 == 3 ~ "Yes",
    TRUE ~ NA_character_  
  )) %>%
  mutate(
    unemployed = fct_recode(unemployed,
                            "No" = "No",
                            "Yes" = "Yes"))

Disc_clean$unemployed <-  ifelse(Disc_clean$unemployed == "No", 0, 
                                 ifelse(Disc_clean$unemployed== "Yes", 1, NA))

with(Disc_clean, table(country, unemployed))
```

### Country_level independent variables

### 7. GDP per capita (Independent Variable) 

```{r}
#dataset from World Bank 
library(readr)

GDP <- read_csv("./Data/GDP.csv")
```

Tidy the data 

```{r}
GDP <- GDP |> na.omit() |>  #remove rows with empty spaces 
  rename(Country =`Country Name`,Indicator = `Series Name`, value = `2019 [YR2019]`) 

GDP$value <- as.numeric(GDP$value)

GDP1 <- GDP |> 
  pivot_wider(names_from = "Indicator", values_from =  "value") |> 
  pivot_wider(names_from = `Series Code`, values_from = `GDP per capita (current US$)`: `GDP (current US$)`)|> 
  rename(GDP_per_capita = `GDP per capita (current US$)_NY.GDP.PCAP.CD`, GDP = `GDP (current US$)_NY.GDP.MKTP.CD`) |>
  select(Country,`Country Code`,GDP_per_capita, GDP)

options(scipen = 999)
GDP1[26,1] <- "Slovakia"
GDP1[18,1] <- "Czech Republic"

GDP1 <- GDP1 |> select(Country, GDP_per_capita) |> mutate(GDP_per_capita = round(GDP_per_capita, 0))

```

### 8. Religious data (Independent Variable)

```{r}
install.packages("xlsx")
library(tidyr)
library(dplyr)

religion <- read_excel("Data/Religious_Composition_by_Country_2010-2050.xlsx")
head(religion)

religion <- religion |> filter(Year == 2020)
religion <- religion |> filter(Region == "Europe")
religion <- religion |> select(Country, "All Religions")

religion <- religion |> rename("country" = "Country")

religion$"All Religions" <- as.numeric(gsub(",", "", religion$"All Religions"))
religion <- religion |> rename("all_religion" = "All Religions")

cyprus <- tibble(country = c("Cyprus"), all_religion = c(1041228))
religion <- rbind(religion, cyprus)

religion <- religion |> filter(country != "Vatican City")

print(religion)
```

RAINBOW DATASET

The following dataset is from the "Rainbow index 2019" of ILGA-EUROPE’S annual benchmarking tool, this index ranks 49 European countries by their LGBTI equality laws and policies, taking into account those policies safeguarding equality and human rights of LGBTI poeple (https://www.ilga-europe.org/report/rainbow-europe-2019/). 

1 - national/federal application 
2 - applicable is some regions only 
3 - No applications 


Each country has a score based on the presence or absence of these laws and policies, as well as the comprehensiveness and effectiveness of such measures,the total score at the end of the dataset reflects the overall level of LGBTQI+ equality in each country (the higher score the better LGBTQI+ equality within the country). 

### 9. Rainbow (Independent variable)

```{r}
library(readxl)
rainbow_2019 <- read_excel("Data/rainbow-2019.xlsx")

head(rainbow_2019)
rainbow_2019 <- rainbow_2019 |> rename("country" = "Country")

rainbow_2019 <- rainbow_2019 |> select(country,"Constitution (gender identity) ()",
                                       "Employment (gender identity) ()", "Education (gender identity) ()",
                                       "LGBTI human rights defenders are not at risk ()", "Score")

rainbow_2019 <- rainbow_2019 |> rename("constitution_rb" = "Constitution (gender identity) ()",
                                       "employment_rb" = "Employment (gender identity) ()",
                                       "education_rb" = "Education (gender identity) ()",
                                       "defenders_at_risk_rb" = "LGBTI human rights defenders are not at risk ()",
                                       "score_rb" = "Score")

```

```{r}
#Top 5 countries with the lowest score of LGBTQI+ equality
rainbow_2019 |> 
  arrange(`score_rb`) |> 
  head(5)
```

### 10. Unemployment (Independent variable)

```{r}
library(dplyr)
unemployment_cntry <- read.csv("Data/API_SL.UEM.TOTL.ZS_DS2_en_csv_v2_80.csv", skip = 4, header = TRUE)
unemployment_cntry <- unemployment_cntry |> select("Country.Name", "X2019")

unemployment_cntry <- unemployment_cntry |> rename("country" = "Country.Name")
unemployment_cntry <- unemployment_cntry |> rename("unemployement_rate" = "X2019")
unemployment_cntry$country <- gsub("Czechia", "Czech Republic", unemployment_cntry$country)
unemployment_cntry$country <- gsub("Slovak Republic", "Slovakia", unemployment_cntry$country)
```

### 11. Political Regimes (Independent variable)

```{r}
library(readr)
political_regime <- read_csv("Data/political-regime.csv")
```

```{r}
library(dplyr)
political_regime <- political_regime |> rename("Political_regime" = "Political regime")
political_regime1 <- political_regime %>%
  filter(Year == "2019" & 
         Entity %in% c("Belgium", "Denmark", "Greece", "Spain", "Finland", "France", 
                       "Ireland", "Italy", "Luxembourg", "Netherlands", "Austria", 
                       "Portugal", "Sweden", "Germany", "United Kingdom", "Bulgaria", 
                       "Cyprus", "Czechia", "Estonia", "Hungary", "Latvia", 
                       "Lithuania", "Malta", "Poland", "Romania", "Slovakia", 
                       "Slovenia", "Croatia"))
```

```{r}
political_regime1 <- political_regime1 %>% 
  mutate(pol_regimen = case_when(
    `Political_regime` == 1 ~ "electoral autocracies",
    `Political_regime` == 2 ~ "electoral democracies", 
    `Political_regime` == 3 ~ "liberal democracies",
    TRUE ~ NA_character_  
  )) %>%
  mutate(pol_regimen = factor(pol_regimen,
                              levels = c("electoral autocracies", "electoral democracies", "liberal democracies")))

political_regime_c <- political_regime1 %>%
  select(Entity, `Political_regime`)


```


## Merging data

```{r}
# select(GDP, GDP_per_capita, country)
survey <- left_join(Disc_clean, GDP1, by = c("country" = "Country")) 
# merge religion data
survey <- left_join(survey, religion, by=c("country" = "country"))
# merge rainbow_2019
survey <- left_join(survey, rainbow_2019, by=c("country" = "country"))
# merge unemployment
survey <- left_join(survey, unemployment_cntry, by=c("country" = "country"))
# merge political regime
survey <- left_join(survey, political_regime_c, by = c("country" = "Entity"))

survey$change_docs <- ifelse(Disc_clean$change_docs == "Yes", 1,
                             ifelse(Disc_clean$change_docs == "No", 0, NA))
survey <- survey |> mutate(across(c(d10, d11, d8r1, d1r1, sd3, sd1_4, sd1_7, sd1_8, qc19,
                           qc20, d70, d15a_r1, age_cat, Life_satisf, left_right, religion,
                           religion_group, contact_gay, contact_trans, contact_inter), as.numeric))

print(survey)
```

## Measuring supprotiveness for LGBTI communities in different countries

Count the qc19 answers by country 

```{r}
library(dplyr)

#sum counts 
opinions_by_country <- Disc_clean %>%
  select(country, change_docs) |> 
  group_by(country, change_docs) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = change_docs, values_from = count, values_fill = 0)

opinions_by_country

#percentage counts 
country_yes_no_dk <- opinions_by_country %>%
  group_by(country) %>%
  mutate(total_count = (Yes+No+DK)) |> 
  mutate(percentage_Yes = (Yes / total_count) * 100) %>%
  mutate(percentage_No = (No / total_count)*100) |> 
  mutate(percentage_DK = (DK / total_count)*100) |> 
  ungroup() |> 
  select(country, percentage_Yes, percentage_No, percentage_DK) 
```


```{r}
#Top 5 of countries that think that transgender persons should be able to change their civil documents to match their inner gender identity
country_with_most_yes <- country_yes_no_dk %>%
  select(country, percentage_Yes) |> 
  filter(percentage_Yes > 0) %>%
  slice_max(percentage_Yes, n=5)

country_with_most_yes
```

```{r}
#Top 5 of countries that do not think that transgender persons should be able to change their civil documents to match their inner gender identity 
country_with_most_no <- country_yes_no_dk %>%
  select(country, percentage_No) |> 
  filter(percentage_No > 0) %>%
  slice_max(percentage_No, n=5)

country_with_most_no
```

```{r}
#Top 5 of countries that do not think that transgender persons should be able to change their civil documents to match their inner gender identity
country_with_most_DK <- country_yes_no_dk %>%
  select(country, percentage_DK) |> 
  filter(percentage_DK > 0) %>%
  slice_max(percentage_DK, n=5)
  
country_with_most_DK
```

### Life Satisfaction (d70)

```{r}
#Life satisfaction by country sum counts 
ls_by_country <- Disc_clean %>%
  select(country, Life_satisf) |> 
  group_by(country, Life_satisf) %>%
  mutate(count = n()) %>%
  ungroup() |> 
  distinct(country, Life_satisf, .keep_all = TRUE) |> 
  pivot_wider(names_from = Life_satisf, values_from = count, values_fill = 0)

ls_by_country
```

### Life Satisfaction percentage

```{r}
#percentage counts 
ls_by_country_pct <- Disc_clean %>%
  group_by(country, Life_satisf) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  group_by(country) %>%
  mutate(total_count = sum(count)) %>%
  mutate(percentage = (count / total_count) * 100) %>%
  select(country, Life_satisf, percentage) %>%
  pivot_wider(names_from = Life_satisf, values_from = percentage, values_fill = 0)

ls_by_country_pct
```



# Missing data imputation

```{r}
sapply(survey, function(x) sum(is.na(x))*100/nrow(survey))

# numerical variable
survey_n = survey |> select(-country, -isocntry)

# we separated two dataset since the scale of data in those two dataset are too different
survey_n_simple <- survey_n |> select(constitution_rb, employment_rb, education_rb, defenders_at_risk_rb, left_right,
                                      religion, religion_group, score_rb, Political_regime)
survey_n_dummy <- survey_n |> select(unemployed, contact_gay, contact_trans, contact_inter, change_docs)

md.pattern(survey_n_simple)
md.pattern(survey_n_dummy)

m = 4 # number of multiple imputations, we are going to make 5 iterations, we're going to predict missing values 4 times.
# survey_n_simple
mice_mod_s = mice(survey_n_simple, m = m, method='cart', seed = 123) # machine learning tool, rf = random forest
survey_n_s_imputed <- complete(mice_mod_s, action=m) 

colSums(is.na(survey_n_s_imputed))

# survey_n_dummy
mice_mod_d = mice(survey_n_dummy, m = m, method='cart', seed = 234) # machine learning tool, rf = random forest
survey_n_d_imputed <- complete(mice_mod_d, action=m) 

colSums(is.na(survey_n_d_imputed))
```

## Data frame

```{r}
survey <- survey |> select(-unemployed, -contact_gay, -contact_trans, -contact_inter, -change_docs,
                 -constitution_rb, -employment_rb, -education_rb, -defenders_at_risk_rb, -left_right,
                -religion, -religion_group, -score_rb, -Political_regime)
survey_imputed <- cbind(survey, survey_n_s_imputed, survey_n_d_imputed)

colSums(is.na(survey_imputed))

# only numberical variable
survey_n_imputed <- survey_imputed |> select(-country, -isocntry)

library(corrplot)
corrplot(cor(survey_n_imputed), method = "number")
```

### Final dataset

```{r}
survey_imputed <- survey_imputed |> select(-d10, -d11, -d8r1, -d1r1, -sd3,
                                           -sd1_4, -sd1_7, -sd1_8, -qc19,  
                                           -qc20, -d70, -d15a_r1)

# only numberical variable
survey_n_imputed <- survey_imputed |> select(-country, -isocntry)
```


# GOALS

## 1. Cross-country differences in supporting levels:

### Logistic Regression

```{r}
LR <- glm(change_docs ~. , data = survey_n_imputed, family = "binomial")
summary(LR)
```

### Logistic regression (country)

```{r}
m0b <- glm(change_docs ~ male + age_cat + I(age_cat^2) + I(GDP_per_capita^2) + I(contact_trans^2) + I(unemployed^2) +
           I(education_rb^2) + factor(country), survey_imputed, family=binomial(link = "logit"))
summary(m0b)
```

## 2. Predictive model for other countries:

### Defining Yes

```{r}
Disc_clean$change_docs <- ifelse(Disc_clean$change_docs == "Yes", 1,
                             ifelse(Disc_clean$change_docs == "No", 0, NA))



survey$agree = factor(survey$change_docs>0) # less than 5 over 20
survey$change_docs = NULL

prop.table(table(survey$agree))
```

### Data splitting 

```{r}
library(caret)

set.seed(123)
in_train <- createDataPartition(survey$agree, p = 0.8, list = FALSE)  # 70% for training
training <- survey[in_train,]
testing <- survey[-in_train,]
nrow(training)
nrow(testing)
length(survey$agree)
```

### 10 fold validation

```{r}
ctrl <- trainControl(method = "repeatedcv", 
                     number = 10,
                     classProbs = T,
                     summaryFunction=twoClassSummary,
                     verboseIter = T)

levels(training$Fail)=c("No","Yes")
levels(testing$Fail)=c("No","Yes")
```

```{r}
# Define a grid for the hyper-parameters
param_grid = expand.grid(gamma = seq(0, 1, 0.1), lambda = seq(0.1, 0.9, 0.1))

# Train to maximize AUC: metric="ROC"
ldaFit <- train(Fail ~ ., 
                method ="rda", 
                data = training,
                tuneGrid = param_grid,
                preProcess = c("center", "scale"),
                metric="ROC",
                trControl = ctrl)
print(ldaFit)

# Predict and validate
ldaPred = predict(ldaFit, testing)
confusionMatrix(ldaPred, testing$Fail)
```

### ROC curve

```{r}
library(pROC)

bench.model = glm(Fail ~ 1, family=binomial(link='logit'), data=training)
prob.bench = predict(bench.model, newdata=testing, type="response")

roc.lda=roc(testing$Fail ~ ldaProb[,2])
roc.bench=roc(testing$Fail ~ prob.bench)

plot(roc.lda, col="red",print.thres=TRUE)
plot(roc.bench, add=TRUE, col='green',print.thres=TRUE)
legend("bottomright", legend=c("lda", "bench"), col=c("red",  "green"), lwd=2)

roc.lda$auc
roc.bench$auc

```

### 1. KNN

```{r}
knnFit <- train(Fail ~ ., 
                  data = training,
                  method = "kknn",   
                  preProc=c('scale','center'),
                  tuneLength = 10,
                  metric="ROC",
                  trControl = ctrl)
plot(knnFit)

knnProb = predict(knnFit, testing, type="prob")
prediction <- as.factor(ifelse(knnProb[,2] > 0.1, "Yes", "No"))

confusionMatrix(prediction, testing$Fail)$table
confusionMatrix(prediction, testing$Fail)$overall[1:2]
```

# Decision Tree (rpart)

```{r}
library(rpart)

control = rpart.control(minsplit = 30, maxdepth = 10, cp=0.01)
```

```{r}
model = agree ~.
dtFit <- rpart(model, data=training, method = "class", control = control)
summary(dtFit)
```

```{r}
library(rpart.plot)
rpart.plot(dtFit, digits=3)
```

```{r}
control = rpart.control(minsplit = 8, maxdepth = 12, cp=0.001)
# Decision Tree Classification
dtFit <- rpart(model, data=training, method = "class", control = control)

rpart.plot(dtFit, digits = 3)
```

# Prediction - Decision Tree

```{r}
dtPred <- predict(dtFit, testing, type = "class")

dtProb <- predict(dtFit, testing, type = "prob")

prediction <- as.factor(ifelse(dtProb[,2] > 0.5, "Yes", "No"))
levels(testing$agree)

confusionMatrix(prediction, testing$agree)$table
confusionMatrix(prediction, testing$agree)$overall[1:2]
```

## Challenges:

In the encoding process, some variables like education () were difficult to encode as we were interested in the level of education, not the age when the participant stopped full-time studies (which does not imply neccesarily the level of studies acquired). A correlation with the ISCED levels could be possible but not rigorous (e.g. older people obtaining their highschool diploma).


